\chapter{Conclusion}
\label{ch:conclusion}

This project proposed an algorithm for disambiguating the sense of words in the
Portuguese language without the need to manually create a sense inventory or
sense-tagged corpora. An experimental implementation was created, which used a
graph-based algorithm and the dependencies produced by \ac{STRING}. And with
this experimental implementation, a database was populated with the
co-occurrences of the CETEMPúblico \textit{corpus}.

Finally, the use of syntactic dependencies for the task of \ac{WSI} was
evaluated and compared with a baseline of the \ac{MFS}.

Future investigation possibilities could focus on improving the results obtained
in this implementation as well as improving the system's performance.

One possibility to improve the results would be to evaluate the use of relaxed
co-occurrences based on context windows, sentences or even paragraphs, instead
of the syntactic dependency co-occurrences used in this project. This might
influence the characteristics of the generated graphs and improve the
performance of the graph-clustering algorithms.

Another aspect should be to investigate vector-based algorithms. These might be
capable of making use of the dependency information provided by \ac{XIP}, and
might perform better in relation to algorithms blind to this
information, such as the ones used in this project.

To improve performance, it might be relevant to investigate porting the data to
a graph database engine. The traditional relational database used requires the
use of multiple \texttt{JOIN}s to execute the required queries. A database with
a native concept of a graph would allow to remove these \texttt{JOIN}s and
improve execution speed.

Although the project was unable to achieve satisfactory results in the
evaluation, it was possible to pinpoint where possible weaknesses are located.
A base model was defined and documented, and it can be used as a base for future
attempts at solving the challenges found.

% kate: default-dictionary en_GB; indent-width 2; replace-tabs on;
% kate: remove-trailing-space on; space-indent on;
% kate: replace-trailing-space-save on; remove-trailing-space on;
